---
title: "Candy Data Cleaning"
author: "John Reyner"
date: "August 23, 2019"
output: html_document
---

```{r}
library(readr)

candyhierarchy2017 <- read_csv("~/candy/candyhierarchy2017.csv")
```

# remove unnecessary columns

```{r}
candy1 <- subset(candyhierarchy2017, select = c(7:109))
```

# change candy reactions to numeric

```{r}
candy1[,1:103] <- ifelse(candy1[,1:103] == "DESPAIR", -1, ifelse(candy1[,1:103] == "MEH", 0, ifelse(candy1[,1:103] == "JOY", 1, 0)))
candy1[is.na(candy1)] <- 0 # it is assumed that any NA value will be a neutral reaction
candy2 <- sapply(candy1, as.numeric)
```

# take means of all candy reactions

```{r}
mean(candy2[,1:103])
median(candy2[,1:103])
```

## this displays a the average reaction to all responses is neutral, however it does not account for non-candy objects

```{r}
positive <- data.frame(react = colMeans(x = candy2) > 0)
is.na(positive) <- positive == "FALSE"
positive <- t(positive)
```

# remove any candy that fails to meet a mean positive reaction

## bind the data frames

```{r}
candy3 <- rbind(candy2, positive)
```

## a potential data set if one wants only positive reactions

```{r}
candy4 <- t(na.omit(t(candy3)))
candy4 <- candy4[-2461,]
```

# remove unidentifiable or non-candy columns
## also there are many different M&M's categories (color) to be dropped in favor of only regular and peanut M&Ms
## the final data set will include all candies

```{r}
candy5 <- subset(candy3, select = -c(2, 3, 6, 8, 9, 13, 15, 16, 20, 21, 24, 25, 26, 32, 37, 38, 
                                     39, 42, 43, 59, 60, 61, 62, 63, 64, 73, 75, 80, 86, 96, 98, 
                                     99, 101, 102))
candy5 <- candy5[-c(2461),]
```

# make set of cleaned data

```{r}
candy <- data.frame(sum = colSums(x = candy5), likes = colSums(candy5 == 1), dislikes = colSums(candy5 == -1), neutral = colSums(candy5 == 0), mean = colMeans(x = candy5), median = apply(candy5, 2, median))
```

# clean up row names
## when plotting the graph, some of the candy names also are very long which makes margining difficult
## those candies will have their names shortened

```{r}
row.names(candy) <- c("100 Grand Bar", "Black Jacks", "Bonkers", "Bottle Caps", "Butterfinger", 
                      "Cadbury Eggs", "Candy Corn", "Caramellos", "Chick-o-Sticks", "Chiclets", 
                      "Coffee Crisp", "Dots", "Dove Bars", "Goo Goo Clusters", "Good N' Plenty", 
                      "Gum", "Gummi Bears", "Hard Candy", "Heath Bar", "Hershey's (Dark)", 
                      "Hershey's (Milk)", "Hershey's Kisses", "JoyJoy", "Junior Mints", 
                      "Kinder Hippo", "Kit Kat", "Laffy Taffy", "Lemon Heads", "Licorice", 
                      "Black Licorice", "Lindt Truffle", "Lollipops", "Mars", "Maynards", 
                      "Mike and Ike", "Milk Duds", "Milky Way", "Regular M&Ms", "Peanut M&Ms", 
                      "Mint Kisses", "Mint Juleps", "Mr. Goodbar", "Necco Wafers", "Nerds", 
                      "Nestle Crunch", "Now'n'Laters", "Peeps", "Pixy Stix", "Reese's Cups", 
                      "Reese's Pieces", "Reggie Jackson", "Rolos", "Skittles", "Smarties (US)", 
                      "Smarties (CW)", "Snickers", "Sourpatch Kids", "Starburst", "SweetTarts", 
                      "Swedish Fish", "Sweetums", "Take 5", "Tic Tacs", "Circus Peanuts", 
                      "Three Musketeers", "Toblerone", "Twix", "Whatchamacallit", "York")
```

# create bar plot for visualization

## segment data for plotting

```{r}
subset <- t(data.frame(candy$likes, candy$dislikes, candy$neutral))
colnames(subset) <- rownames(candy)
rownames(subset) <- c("likes", "dislikes", "neutral")
```

## plot data

```{r}
par(mar = c(9, 4, 4, 6))
barplot(subset, names.arg = candy$Input, main = "Candy Favorability Levels", 
        ylab = "Responses", beside = FALSE, col = c("green", "brown", "gray"), 
        legend = c("Likes", "Dislikes", "Neutral"), 
        args.legend = list(x = ncol(subset) + 30, bty = "n"),
        las = 2)
```

## create a bar plot of favorite candies in ascending order

```{r}
ss2 <- as.data.frame(t(subset))
fav1 <- as.matrix(t(ss2[with(ss2, order(likes)),]))

par(mar = c(9, 4, 4, 6))
barplot(fav1, names.arg = candy$Input, main = "Candy Favorability Levels", 
        ylab = "Responses", beside = FALSE, col = c("green", "brown", "gray"), 
        legend = c("Likes", "Dislikes", "Neutral"), 
        args.legend = list(x = ncol(subset) + 30, bty = "n"),
        las = 2)
```
